### 2022-03-08 01-processing-raw-data_reviews.Rmd
first commit

---
title: "Processing of raw data"
author: "Ulrik Lyngs"
date: '2022-03-01'
output: html_document
editor_options: 
  chunk_output_type: inline
---

```{r}
library(tidyverse)
library(rvest)
library(janitor)
library(jsonlite)
```

# Chrome Web store
## Extract reviews from page sources
In this case, we've store .RDS files that contain the page sources.

We'll first create the helper functions we need.

```{r}
get_reviews_from_full_source <- function(source) {
  source %>% 
    read_html() %>% 
    html_nodes(".ba-fb > div > .ba-bc-Xb")
}

extract_info_from_review_div <- function(review_div) {
  user_name <- review_div %>% 
    html_node(".comment-thread-displayname") %>% 
    html_text()

  date_modified <- review_div %>% 
    html_node(".ba-Eb-Nf") %>% 
    html_text()

  numstars <- review_div %>% 
    html_nodes(".rsw-starred") %>% 
    length()

  text <- review_div %>% 
    html_node(".ba-Eb-ba") %>% 
    html_text()

  result_tibble <- tibble(type = "review",
                          user_name,
                          reply_to = NA,
                          date_modified,
                          numstars,
                          text)

  replies <- review_div %>% 
    html_nodes(".Fg-b-ob-fb .ba-bc-Xb-K")

  if (length(replies) > 0) {
    for (i in seq_along(replies)) {
      name_replier <- replies[[i]] %>% 
        html_node(".comment-thread-displayname") %>% 
        html_text()

      date_modified <- replies[[i]] %>% 
        html_node(".ba-Eb-Nf") %>% 
        html_text()

      text <- replies[[i]] %>% 
        html_node(".ba-Eb-ba") %>% 
        html_text()

      result_tibble <- bind_rows(result_tibble,
                                 tibble(type = "reply",
                                        reply_to = user_name,
                                        user_name = name_replier,
                                        date_modified,
                                        numstars = NA,
                                        text))
    }
  }

  return(result_tibble)

}

get_review_content_from_list <- function(list_with_review_divs) {
  map_dfr(list_with_review_divs, extract_info_from_review_div)
}

return_reviews_from_source <- function(tibble_with_page_sources) {
  tibble_with_page_sources %>% 
    mutate(review_divs = map(unlist(value), get_reviews_from_full_source),
           review_content = map(review_divs, get_review_content_from_list)) %>% 
    select(-value, -review_divs) %>% 
    rename(review_page = name) %>% 
    unnest(cols = c(review_content)) %>% 
    filter(!is.na(text))
}
```

We'll do things in two steps, to avoid running into memory issues.

We first just read in the RDS files one at a time, extract the reviews, and store them in a CSV file (note: we're not doing this in a very efficient way, so it might take dozens of seconds for extensions with many reviews, e.g. 'Block Site'):

It takes a while, so we'll set it to not evaluate by default.

```{r eval=FALSE}
#dir.create(here::here("data", "processed", "chrome", "2019-02-08-reviews_extracted"))

read_process_and_store <- function(filepath, filename) {
  savename <- filename %>% str_remove("\\.rds") %>% make_clean_names()

  print(str_c("doing ", filename))

  read_rds(filepath) %>% 
    return_reviews_from_source() %>% 
    write_csv(str_c(here::here("data", "processed", "chrome", "2019-02-08-reviews_extracted/"), savename, ".csv"))
}

rds_files_w_reviews <- tibble(
  filepath = list.files(path = here::here("data", "raw", "chrome", "2019-02-08-page_sources"), pattern = "rds$", full.names = TRUE),
  filename = basename(filepath)
)

rds_files_w_reviews %>% 
  pwalk(read_process_and_store)
```

## Clean up and save out

```{r}
review_CSVs <- list.files(path = here::here("data", "processed", "chrome", "2019-02-08-reviews_extracted"), pattern = "csv$", full.names = TRUE)

chrome_reviews <- map_dfr(review_CSVs, read_csv, 
                          col_types = cols(
                            numstars = col_integer(),
                            reply_to = col_character())
                          )

chrome_reviews %>% 
  # add date
  mutate(date_modified = str_remove(date_modified, "Modified "),
         date_modified = case_when(
          str_detect(date_modified, "hours ago") ~ "Feb 8, 2019",
          str_detect(date_modified, "1 day ago") ~ "Feb 7, 2019",
          str_detect(date_modified, "2 days ago") ~ "Feb 6, 2019",
          str_detect(date_modified, "3 days ago") ~ "Feb 5, 2019",
          str_detect(date_modified, "4 days ago") ~ "Feb 4, 2019",
          str_detect(date_modified, "5 days ago") ~ "Feb 3, 2019",
          str_detect(date_modified, "6 days ago") ~ "Feb 2, 2019",
          TRUE ~ date_modified
        ),
    date_modified = lubridate::mdy(date_modified)) %>% 
  rename(rating = numstars) %>% 
  distinct(across(-review_page)) %>% 
  write_csv(here::here("data", "processed", "chrome", "2019-02-08-chrome_reviews_unique.csv"))

```

# Google Play
Here we have all the reviews in one big json file

```{r}
play_reviews_json <- read_json(here::here("data", "raw", "play", "2019_03_19_play_all_reviews.json"))

# have a look at the first app
# play_reviews_json[[1]] %>% 
#   View()

# put the data in a tibble
play_reviews <- tibble(
  app_id = map_chr(play_reviews_json, "packageName"),
  num_reviews = map_int(play_reviews_json, "numberOfReviews"),
  reviews = map(play_reviews_json, "reviews")
) %>% 
  arrange(desc(num_reviews))

# let's just see how many we have
play_reviews %>% 
  summarise(num_apps = n(),
    num_reviews = sum(num_reviews))

```

Now we grab the information we want from the reviews and put it in a tibble

```{r}
# let's make a function to extract the information we want from the reviews
extract_play_review_info <- function(play_store_review) {
  tibble(
    user_name = map_chr(play_store_review, "userName"),
    date = map_chr(play_store_review, "date"),
    rating = map_int(play_store_review, "score"),
    text = map_chr(play_store_review, "text")
  )
}

# put 'em all in one big happy unnested tibble
play_reviews_tibble <- play_reviews %>% 
  mutate(review_info = map(reviews, extract_play_review_info)) %>% 
  select(-reviews, -num_reviews) %>% 
  unnest(cols = c(review_info))

```

## Save out unique reviews
Then we remove duplicates and save them out

```{r}
play_reviews_tibble %>% 
  mutate(date = lubridate::mdy(date)) %>% 
  distinct() %>% 
  write_csv(here::here("data", "processed", "play", "2019_03_19-play_reviews_unique.csv"))

```

# Apple App Store
Here we have things in separate files.

## JSON files from the GitHub scraper
The scraper from https://github.com/facundoolano/app-store-scraper has given us JSON files

```{r}
# have a look at one
(example_apple_json <- read_json(here::here("data", "raw", "apple", "gh_scraper_2019-03-19_individual_reviews", "club.donutdog.ios.json")))

# grab filepaths
json_files <- list.files(path  = here::here("data", "raw", "apple", "gh_scraper_2019-03-19_individual_reviews"), pattern = "json$", full.names = TRUE)

apple_reviews_json <- json_files %>% 
  map(read_json)

apple_reviews <- tibble(
  app_id = map_chr(apple_reviews_json, "app"),
  reviews = map(apple_reviews_json, "reviews"),
  num_reviews = map_int(reviews, length)
)

# let's just see how many we have
apple_reviews %>% 
  summarise(num_apps = n(),
            num_reviews = sum(num_reviews))

```

Let's actually grab the information we want and put in a useful format.

NOTE: Apple reviews have a heading for the reviews - we'll merge this into the text.

```{r}
extract_apple_review_info <- function(apple_review) {
  tibble(
    user_name = map_chr(apple_review, "userName"),
    rating = map_int(apple_review, "score"),
    review_title = map_chr(apple_review, "title"),
    review_text = map_chr(apple_review, "text")
  )
}

# put 'em all in one big happy unnested tibble
apple_review_tibble <- apple_reviews %>% 
  mutate(review_info = map(reviews, extract_apple_review_info)) %>% 
  select(-reviews) %>% 
  unnest()

```

We noticed on manual inspection that these apps had reviews on their store pages that the github scraper missed:
- pabloweb.net.SelfControl
- com.imobiapp.screentimer
- com.getcluster.Compose
- club.donutdog.ios
- com.getfeedless.feedless
- com.riko.suyasaso
- com.bellostudios.hooked
- com.jordan-carney.Liberate
- org.aliyilmaz.notelr
- ca.genoe.Refrain
- vn.wehelp.SelfControlLite
- com.yeowjin.simpletext
- com.usedopamine.app.space
- com.erichuju.pomodoro
- com.bluecocoa.time2

We scraped the page sources for these apps ourselves. Let's add them in.

## Extract reviews from page sources
```{r}
apps_w_missing_reviews <- c('pabloweb.net.SelfControl','com.imobiapp.screentimer','com.getcluster.Compose','club.donutdog.ios','com.getfeedless.feedless','com.riko.suyasaso','com.bellostudios.hooked','com.jordan-carney.Liberate','org.aliyilmaz.notelr','ca.genoe.Refrain','vn.wehelp.SelfControlLite','com.yeowjin.simpletext','com.usedopamine.app.space','com.erichuju.pomodoro','com.bluecocoa.time2')

length(apps_w_missing_reviews)
```

Let's make the helper functions:

```{r}
# for extracting developer responses
extract_dev_response_text <- function(review_div){
  html_nodes(review_div, '.we-customer-review__body') %>% .[[2]] %>%  html_text()
}
safe_extract_dev_response_text <- possibly(extract_dev_response_text, NA_character_)

grab_reviews_from_single_div <- function(single_review_div){
  tibble(
    user_name = html_node(single_review_div, '.we-customer-review__user') %>% html_text(),
    date = html_node(single_review_div, 'time') %>% html_text(),
    rating = html_node(single_review_div, '.we-customer-review__rating') %>% html_attrs() %>% enframe() %>% filter(name == 'aria-label') %>% pull(value),
    review_title = html_node(single_review_div, '.we-customer-review__title') %>% html_text(),
    review_text = html_node(single_review_div, '.we-customer-review__body') %>% html_text(),
    dev_response_title = html_node(single_review_div, '.we-customer-review__header--response .we-customer-review__title') %>% html_text(),
    dev_response_date = html_node(single_review_div, '.we-customer-review__header--response time') %>% html_text(),
    dev_response_text = safe_extract_dev_response_text(single_review_div)
  )

}

iterate_over_review_divs <- function(div_with_reviews){
  map_dfr(div_with_reviews, grab_reviews_from_single_div)
}

```

Then let's do it

```{r}
reviews_visible_on_store_pages <- tibble(
  app_id = apps_w_missing_reviews,
  html = map(apps_w_missing_reviews, ~ read_html(paste0(here::here("data", "raw", "apple", "own_scraper_2019-03-20_page_sources/"), .x, ".xml"))),
  review_divs = map(html, ~ html_nodes(.x, ".l-row .we-customer-review")),
  reviews = map(review_divs, iterate_over_review_divs)
) %>% 
  select(app_id, reviews) %>% 
  unnest(cols = c(reviews))

```

## Merge reviews w/ github scraper
Finally, we take reviews from what was scraped with the ordinary scraper from github, and merge with those we scraped ourselves.

```{r}
apple_reviews_all <- reviews_visible_on_store_pages %>% 
  mutate(rating = str_extract(rating, "\\d") %>% as.integer()) %>% 
  bind_rows(apple_review_tibble %>% select(-num_reviews))

```

## Clean up and save out unique reviews

```{r}
apple_reviews_all %>% 
  # merge the title of the apple reviews into the text, then remove line breaks
  mutate(user_name = str_remove(user_name, "\n"),
         review_text = str_c(review_title, review_text, sep = ". ") %>% str_replace_all("\n", "")) %>% 
  select(-review_title) %>% 
  # turn the rating into an integer
  mutate(rating = str_replace(rating, "out of 5", "") %>% as.integer()) %>% 
  # change date format
  mutate(date = lubridate::dmy(date)) %>% 
  distinct() %>% 
  write_csv(here::here("data", "processed", "apple", "2019-03-21_apple_reviews_unique.csv"))
```