### 2022-03-08 03-wrangling_metadata.Rmd

---
title: "Wrangling - meta data"
author: "Ulrik Lyngs"
date: '2022-03-04'
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
library(tidyverse)
library(jsonlite)
```

# Combine the meta data
## Read in
### Chrome & Apple
These are straight-forward for us to read, we simply do it:

```{r, message=FALSE}
# chrome
chrome_tools <- read_csv(here::here("data", "processed", "chrome", "2019-03-21-chrome_meta_data.csv")) %>% 
  mutate(store = "chrome")

apple_tools <- read_csv(here::here("data", "processed", "apple", "2019-03-20-app_info_extracted_from_page_sources.csv")) %>% 
  mutate(store = "apple")
```

### Play
From the below, the apps that were unavailable will have 'null' as their title. 
We'll exclude them, then afterwards we'll unlist the list columns.

```{r}
# play store
play_tools_raw <- read_json(here::here("data", "raw", "play", "2019-03-19_play_app_info.json"))

play_tools_w_nulls <- tibble(
  app = map_chr(play_tools_raw, "app"),
  app_header_title = map(play_tools_raw, "results") %>% map("title"), #note, is null if couldn't scrape
  app_header_subtitle = NA,
  price = map(play_tools_raw, "results") %>% map("price"),
  company = map(play_tools_raw, "results") %>% map("developer"),
  rating_and_count = NA,
  summary = map(play_tools_raw, "results") %>% map("summary"),
  description = map(play_tools_raw, "results") %>% map("description"),
  version = map(play_tools_raw, "results") %>% map("version"),
  average_rating = map(play_tools_raw, "results") %>% map("score"),
  num_ratings = map(play_tools_raw, "results") %>% map("ratings"),
  num_reviews = map(play_tools_raw, "results") %>% map("reviews"),
  num_5_star_reviews = map(play_tools_raw, "results") %>% map("histogram") %>% map("5"),
  num_4_star_reviews = map(play_tools_raw, "results") %>% map("histogram") %>% map("4"),
  num_3_star_reviews = map(play_tools_raw, "results") %>% map("histogram") %>% map("3"),
  num_2_star_reviews = map(play_tools_raw, "results") %>% map("histogram") %>% map("2"),
  num_1_star_reviews = map(play_tools_raw, "results") %>% map("histogram") %>% map("1"),
  min_installs = map(play_tools_raw, "results") %>% map("minInstalls")
)

play_tools <- play_tools_w_nulls %>% 
  filter(!map_lgl(app_header_title, is.null)) %>%  # because app_header_title is a list column, we need to use map
  mutate_if(is.list, unlist) %>% 
  mutate(store = "play")
```

## Combine
```{r}
apple_merge <- apple_tools %>% 
  mutate(num_reviews = NA_integer_,
         min_installs = NA_integer_,
         num_users = NA_integer_) %>% 
  select(store, tool_id = app, title = app_header_title, company, average_rating, num_ratings, num_reviews, min_installs, num_users)

play_merge <- play_tools %>% 
  mutate(num_users = NA_integer_) %>% 
  select(store, tool_id = app, title = app_header_title, company, average_rating, num_ratings, num_reviews, min_installs, num_users)

chrome_merge <- chrome_tools %>% 
  mutate(company = str_replace(company, "offered by", "")) %>% 
  # use the unique identifier in the URL as tool id
  mutate(tool_id = str_extract(extension_url, "/[^/]+$") %>% str_remove("/")) %>% 
  mutate(num_reviews = NA_integer_,
         min_installs = NA_integer_) %>% 
  select(store, tool_id, title = extension_header_title, company, average_rating, num_ratings, num_reviews, min_installs, num_users, company)

tool_info <- apple_merge %>% 
  bind_rows(play_merge) %>% 
  bind_rows(chrome_merge)

```

## Add Lyngs et al.'s feature categorisation
```{r}
# get the original feature coding from Lyngs et al, and create a column with its main types of design patterns in alphabetical order
original_tools_w_alphabetical_categories <- read_csv(here::here("materials", "2018-12-30_lyngs_tools.csv")) %>% 
  select(tool_id, url, `main category`:`main category 4`) %>% 
  unite(categories, c(`main category`,`main category 2`,`main category 3`,`main category 4`), sep = " + ") %>% 
  mutate(categories = categories %>% str_replace_all(" \\+ NA", "")) %>%   #unite doesn't have a drop.na option
  mutate(categories_alphabetically = str_split(categories, " \\+ ")) %>% 
  rowwise() %>%
  mutate(num_categories = length(categories_alphabetically)) %>%
  mutate(categories_alphabetically = categories_alphabetically %>% str_sort() %>% str_c(collapse = " + ")) %>%
  ungroup() %>% 
  select(-categories)

tool_info %>% 
  left_join(original_tools_w_alphabetical_categories) %>% 
  relocate(title, tool_id, store) %>% 
  write_csv(here::here("data", "processed", "meta_data_all_w_design_pattern_types.csv"))
```