### 2022-03-08 materials/scrapers/chrome-web/chrome_ext_info_scraper.Rmd

---
title: "Chrome extensions info scraper"
author: "Ulrik Lyngs"
date: "21 March 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(rvest)
library(lubridate)
library(here)
library(xml2)
library(janitor)
```

This sript was originally used to scrape tools from the Apple App store on 19 March 2019.

# Scraping with rvest
## Get page sources

### Rescrape
If you want to rescrape the tools from the store, run this

```{r, eval=FALSE}
tools_to_scrape <- read_csv(here::here("materials", "2018-12-30_lyngs_tools.csv")) %>% 
  filter(store == "Chrome Web Store") %>% 
  select(title, url) %>% 
  mutate(filename = str_c("ext", row_number(), "_", make_clean_names(title)))

safe_read <- safely(read_html)

tools_with_html <- tools_to_scrape %>% 
  mutate(html = map(url, safe_read)) %>% # SCRAPE
  mutate(is_ok = map_lgl(html, ~ is_null(.$error))) %>% # add column saying whether the scrape worked
  filter(is_ok == TRUE) %>% # only the keep the ones that were available
  mutate(html = map(html, "result")) %>% # grab just the html part

# create a folder to store the page sources in
dir.create(here("data", "raw", "chrome", str_c(today(), "-page_sources")))

# save them out
tools_with_html %>% 
  select(html, filename) %>% 
  pwalk(~write_xml(.x, str_c(str_c(here("data", "raw", "chrome", str_c(today(), "-page_sources_info/")), .y, ".xml"))))

```

### Use original data
If you want to do this with the original data from March 2019, run this

```{r, warning=FALSE}
tools_with_html <- tibble(
  filename = list.files(here("data", "raw", "chrome", "2019-03-21-page_sources_info")) %>% str_remove("\\.xml$"),
  html_paths = list.files(here("data", "raw", "chrome", "2019-03-21-page_sources_info"), full.names = TRUE)
) %>% 
  mutate(html = map(html_paths, read_html)) %>% 
  select(-html_paths)

```

## Extract app info

```{r}
extract_data <- function(html){
  tibble(
    extension_url = html_node(html, '[property="og:url"]') %>% html_attr("content"),
    extension_header_title = html_node(html, '.e-f-w') %>% html_text(),
    company = html_node(html, '.e-f-Me') %>% html_text(),
    summary = html_node(html, '.C-b-p-j-Pb') %>% html_text(),
    description = html_node(html, '.C-b-p-j-Oa') %>% html_text(),
    version = html_node(html, '.C-b-p-D-Xe.h-C-b-p-D-md') %>% html_text(),
    average_rating = html_node(html, '.q-N-nd') %>% html_attrs() %>% enframe() %>% filter(name == 'aria-label') %>% pull(value),
    num_ratings = html_node(html, '.q-N-nd') %>% html_attrs() %>% enframe() %>% filter(name == 'aria-label') %>% pull(value),
    num_users = html_node(html, '.e-f-ih') %>% html_text()
  )
}

# get the meta data
extension_info <- map_dfr(tools_with_html$html, extract_data)

```

Now let's do a bit of cleaning of the data and save it out

```{r}
info_cleaned <- extension_info %>% 
  mutate(average_rating = str_extract(average_rating, "\\d[^o]+") %>% parse_number(),
         num_ratings = str_extract(num_ratings, "[^\\.]+ user") %>% str_replace("user", "") %>% str_replace("One", "1") %>% parse_number(),
         num_users = str_replace(num_users, "user", "") %>% str_replace("s", "") %>% parse_number(),
         company = str_replace(company, "offered by ", "") %>% str_replace("\n", ""))

info_cleaned %>% 
  write_csv(here("data", "processed", "chrome", "2019-03-21-chrome_meta_data.csv"))

```