### 2022-03-08 materials/scrapers/chrome-web/chrome_review_scraper.Rmd

---
title: "chrome_reviews_scraper"
author: "Ulrik Lyngs"
date: "05/02/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(robotstxt)
library(jsonlite)
library(tidyverse)
library(rvest)
library(RSelenium)
library(here)
library(RCurl)
```

# Rvest <3 RSelenium
## Websites' own stated rules - robots.txt
From a research ethics perspective, we should start by inspecting **robots.txt**, which is a file most websites contain saying how they want others to behave when accessing the website using robots.

```{r}
get_robotstxt(domain = "chrome.google.com/")

paths_allowed("https://chrome.google.com/webstore/detail/")

```

So we see that Google actually don't want us to search through the Chrome Web Store with automated scrapers, but everything else should be fair game.

We also add delays when we crawl so that we're not mistaken for a DDoS attack.

We'll do this with `Sys.sleep()`, e.g. `Sys.sleep(5)` to add a 5-second delay. However, using a constant might also get you recognised as a robot. We could use e.g. `Sys.sleep(runif(n = 1, min = 5, max = 10))` to add a random sleep between 5 and 10 seconds.

## Scraping with RSelenium
### Getting things up and running

1. install RSelenium
```{r}
# to run RSelenium we need these:
# devtools::install_github("johndharrison/binman")
# devtools::install_github("johndharrison/wdman")
# devtools::install_github("ropenSci/RSelenium")
```

2. Get Docker up and running
  - install it from https://docs.docker.com/docker-for-mac/install/
  - Open the app (on first run you will need to accept its terms and conditions)
  - then in a terminal start a chrome server with `docker run --name chrome -d -p 4445:4444 selenium/standalone-chrome`
  - there's some test things shown here that you can run (commented out)

```{r, eval=FALSE}
# in a terminal run: docker run --name chrome -d -p 4445:4444 selenium/standalone-chrome
# this will start up a server named 'chrome'
# to stop it you go 'docker stop chrome'
# to see what's currently running, go 'docker ps'
# to delete a server instance name chrome, do 'docker rm chrome'

# -name names your container
# -v mount volume
# -d detached mode (if you don't do this, it'll actually enter the image rather than just)
# -p port mapping (selenium connects to port 4444 by default; if you go like 4445:4444 then you can connect to 4445 on the outside which is mapped to 4444; with -p 5901:5900 you can open up ports for viewers)

# remDr <- remoteDriver(port=4445L, browserName = "chrome")
# remDr$open()
# remDr$getStatus()
# remDr$navigate("https://www.google.com/")
# remDr$getCurrentUrl()
# remDr$navigate("https://www.bing.com/")
# remDr$getCurrentUrl()
# remDr$goBack()
# remDr$getCurrentUrl()
# remDr$goForward()
# remDr$getCurrentUrl()
# remDr$close()
```

### Create function to scrape the sources of review pages

```{r, echo=TRUE, eval=FALSE, results='hide'}
scrape_review_pages <- function(url_string){
  # navigate to url
  print(str_c("navigating to ", url_string))
  remDr$navigate(url_string)
  Sys.sleep(1) # wait for one second
  print(remDr$getCurrentUrl())
  remDr$screenshot(display = TRUE) # show screenshot

  # click 'reviews' button
  reviewButton <- remDr$findElement(using = 'class', "h-e-f-z-b")
  Sys.sleep(0.5)
  reviewButton$clickElement()
  remDr$screenshot(display = TRUE) # show screenshot

  # save the source
    # we don't know how many reviews there are and because
    # speed is not really a concern, we won't preallocate output length
  n <- 1
  reviews <- list()

  reviews[[n]] <- remDr$getPageSource()

  # while there is a visible 'next' button, click it and save the information
  nextButton <- remDr$findElement(using = 'css', "a.Aa.dc-se")
  Sys.sleep(0.5)
  remDr$executeScript("arguments[0].scrollIntoView(true);", args = list(nextButton)) # scroll to it
  remDr$screenshot(display = TRUE)

  while (nextButton$isElementDisplayed()[[1]]) {
    nextButton$clickElement()
    remDr$screenshot(display = TRUE) # show screenshot

    # store the page
    n <- n + 1
    print(stringr::str_c("storing page ", n))
    reviews[[n]] <- remDr$getPageSource()

    print(str_c("can we see a 'next' button?", nextButton$isElementDisplayed()[[1]]))

    print("sleeping for one second")
    Sys.sleep(1)
  }

  return(reviews)
}
```

### Create function to iterate over list of titles and urls and scrape their sources

```{r}
# make helper function to strip out special characters from titles
remove_special_chars <- function(annoying_title) {
  str_replace_all(annoying_title, "[^[A-Za-z]]", " ") %>% str_replace_all(.,"[ ]+", " ")  
}

# make helper function to name the list with the pages
set_review_page_names <- function(review_pages){
  page_names <- str_c("page ", seq_along(review_pages))

  review_pages %>% 
    set_names(page_names)
}

scrape_and_store_page_sources <- function(title_and_url_tbl) {
  for (i in 1:nrow(title_and_url_tbl)) {
    print(stringr::str_c("there are ", nrow(title_and_url_tbl), " extensions to scrape."))

    print(stringr::str_c("scraping extension number ", i, " titled ", title_and_url_tbl$title[i], " from url ", title_and_url_tbl$url[i]))

    # get the review pages
    review_pages <- scrape_review_pages(title_and_url_tbl$url[i])

    # put the result and its parsing in a tibble
    source_tibble <- review_pages %>% 
      set_review_page_names() %>% 
      enframe() %>% 
      mutate(extension_title = title_and_url_tbl$title[i],
             extension_url = title_and_url_tbl$url[i],
             time_scraped = lubridate::now())

    # store it
    source_tibble %>% 
      write_rds(str_c(here::here("data", "raw", "chrome", str_c(lubridate::today(), "-page_sources_reviews/")), "extension_num_", i, remove_special_chars(title_and_url_tbl$title[i]), ".rds"))
  }
}

```

### Error handling
So we need this to be able to handle errors. Let us try with three extensions - two we know works and one we know doesn't work.

```{r}
# Lights off will give an error - the others won't
test_extensions <- tibble(
  title = c("(NO)tifications", "Lights off", "Afairo"),
  url = c("https://chrome.google.com/webstore/detail/notifications/ocfjkjbhicdjhmgkolejpojieddlogbg?hl=gb", "https://chrome.google.com/webstore/detail/lights-off/ehfaompdnhilblcmkkakofflcjfijhai?hl=gb", "https://chrome.google.com/webstore/detail/afairo/imecoioakflafgnkkklfhopabameijck?hl=gb")
)
```

So rather than running this directly, we'll wrap the scraper function in `safely` to make sure we don't crash when hitting extensions that have become unavailable. See [R4DS 'Dealing with failure'](https://r4ds.had.co.nz/iteration.html#dealing-with-failure).

```{r}
safe_scrape_sources <- safely(scrape_and_store_page_sources)
```

Now let's test it.

```{r}
# TEST: scrape review page sources for forest
forest_reviews <- safe_scrape_sources("https://chrome.google.com/webstore/detail/forest-stay-focused-be-pr/kjacjjdnoddnpbbcjilcajfhhbdhkpgk?hl=gb")
```

### Go ahead and do it!

```{r}
extensions_to_scrape <- read_csv(here("materials", "2018-12-30_lyngs_tool_coding.csv")) %>% 
  filter(is.na(Exclude)) %>% 
  filter(str_detect(url, "chrome.google.com/webstore")) %>% 
  mutate(is_available = url.exists(url)) %>%  # this will take a while
  filter(is_available == TRUE) %>% 
  select(title, url)

scrape_and_store_page_sources(extensions_to_scrape)

```