### 2022-06-10 figures/figures.Rmd
correct figure plot code

---
title: "Figure code"
output: html_document
---

```{r setup, include=FALSE}
library(tidyverse)
library(cowplot)
library(ggtextures) #remotes::install_github("clauswilke/ggtextures") 
```

```{r read_data}
meta_data <- read_csv(here::here("data", "processed", "meta_data_all_w_design_pattern_types.csv"))
original_tools <- read_csv(here::here("materials", "2018-12-30_lyngs_tools.csv"))
```

```{r}
# set up the mapping between categories and colors
tools_w_categories_and_image <- meta_data %>% 
  mutate(image_single = case_when(
    categories_alphabetically == "Block/removal" ~ "1_block",
    categories_alphabetically == "Goal-advancement" ~ "1_goal",
    categories_alphabetically == "Self-tracking" ~ "1_track",
    categories_alphabetically == "Reward/punish" ~ "1_reward",
    categories_alphabetically == "Block/removal + Goal-advancement" ~ "2_block_goal",
    categories_alphabetically == "Block/removal + Reward/punish" ~ "2_block_reward",
    categories_alphabetically == "Block/removal + Self-tracking" ~ "2_block_track",
    categories_alphabetically == "Goal-advancement + Reward/punish" ~ "2_goal_reward",
    categories_alphabetically == "Goal-advancement + Self-tracking" ~ "2_track_goal",
    categories_alphabetically == "Reward/punish + Self-tracking" ~ "2_track_reward",
    categories_alphabetically == "Block/removal + Goal-advancement + Self-tracking" ~ "3_block_track_goal",
    categories_alphabetically == "Block/removal + Reward/punish + Self-tracking" ~ "3_block_track_reward",
    categories_alphabetically == "Goal-advancement + Reward/punish + Self-tracking" ~ "3_goal_track_reward",
    categories_alphabetically == "Block/removal + Goal-advancement + Reward/punish + Self-tracking" ~ "4_all",
  )) %>% 
  mutate(image_single = str_c(here::here("figures", "plot_colours/"), image_single, ".png")) %>% 
  select(tool_id, title, num_ratings, min_installs, num_users, store, image_single)

```

# Plot of tools by user numbers
## generate apple plot
```{r}
apple_data <- tools_w_categories_and_image %>% 
  filter(store == "apple") %>% 
  filter(num_ratings > 0)

images <- apple_data$image_single %>% 
  set_names(apple_data$title)

images_fill = c(
  chrome =  here::here("figures", "textures", "fillpatterns3.gif"), #"https://www.pyngl.ucar.edu/Graphics/Images/fillpatterns3.gif",
  play = here::here("figures", "textures", "siding1-256.jpg"),  #"http://www.hypergridbusiness.com/wp-content/uploads/2012/12/siding1-256.jpg",
  apple = here::here("figures", "textures", "concrete1-256.jpg") # "http://www.hypergridbusiness.com/wp-content/uploads/2012/12/concrete1-256.jpg"
)

apple_features <- apple_data %>% 
  ggplot() +
    geom_textured_bar(aes(x = reorder(title, num_ratings), y = -1000, image = title), stat = "identity", width = 0.8) +
    scale_image_manual(values = images) +
    coord_flip() +
    labs(x = "", y = "") +
    scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
    #scale_y_log10() +
    guides(image = "none") +
    theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank()) +
    theme(plot.margin = margin(0.6, 0, 0.8, 0, "cm"))

apple_bar_chart <- apple_data %>% 
  ggplot() +
    geom_textured_bar(aes(x = reorder(title, num_ratings), y = num_ratings, image = store), stat = "identity", width = 0.8) +
    scale_image_manual(values = images_fill) +
    labs(x = "", y = "Number of ratings (log scale)") +
    scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
    scale_y_log10(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
    guides(image = FALSE) +
    theme_minimal() +
    theme(plot.margin = margin(0.6, 0, 0, 0, "cm")) +
    coord_flip()

apple_combined <- plot_grid(apple_bar_chart, apple_features, rel_widths = c(4,1), labels = c("Apple App", "Features"), hjust = "left", label_x = c(0.02, 0.075), label_y = 1)

save_plot(here::here("figures", "apple_ratings_and_features.png"), apple_combined,
          # each individual subplot should have an aspect ratio of 1.3
          base_aspect_ratio = 1.8
          )
```

## generate google play plot

```{r}
num_tools_play_chrome <- 25

##### create the google play plot
play_data <- tools_w_categories_and_image %>% 
  filter(store == "play") %>% 
  arrange(desc(min_installs)) %>%
  slice(1:num_tools_play_chrome)

##### create the google play plot
images <- play_data$image_single %>% 
  set_names(play_data$title)

play_feature_plot <- play_data %>% 
  ggplot() +
    geom_textured_bar(aes(x = reorder(title, min_installs), y = -1000000, image = title), stat = "identity", width = 0.8) +
    scale_image_manual(values = images) +
    coord_flip() +
    labs(x = "", y = "") +
    scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
    guides(image = FALSE) +
    theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank()) +
    theme(plot.margin = margin(0.6, 0, 0.8, 0, "cm"))

play_install_bar <- play_data %>%
  ggplot() +
    geom_textured_bar(aes(x = reorder(title, min_installs), y = min_installs, image = store), stat = "identity", width = 0.8) +
    scale_image_manual(values = images_fill) +
    labs(x = "", y = "Minimum number of installs (log scale)") +
    scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
    scale_y_log10(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
    guides(image = "none") +
    theme_minimal() +
    theme(plot.margin = margin(0.6, 0, 0, 0, "cm")) +
    coord_flip()

play_combined <- plot_grid(play_install_bar, play_feature_plot, rel_widths = c(4,1), labels = c("Google Play", "Features"), hjust = "left", label_x = c(0.02, 0.075), label_y = 1)

save_plot(here::here("figures", "play_installs_and_functionality.png"), play_combined,
          # each individual subplot should have an aspect ratio of 1.3
          base_aspect_ratio = 1.8
          )
```

## generate chrome web store plot

```{r}
##### create the chrome web plot
chrome_data <- tools_w_categories_and_image %>% 
  filter(store == "chrome") %>% 
  arrange(desc(num_users)) %>% 
  slice(1:num_tools_play_chrome)

##### create the chrome extensions plot
images <- chrome_data$image_single %>% 
  set_names(chrome_data$title)

chrome_feature_plot <- chrome_data %>% 
  ggplot() +
    geom_textured_bar(aes(x = reorder(title, num_users), y = -200000, image = title), stat = "identity", width = 0.8) +
    scale_image_manual(values = images) +
    coord_flip() +
    labs(x = "", y = "") +
    scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
    guides(image = FALSE) +
    theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank()) +
    theme(plot.margin = margin(0.6, 0, 0.8, 0, "cm"))

# IF WE WANT THE FILL TO BE THE SAME AS FOR RATINGS
chrome_user_bar <- chrome_data %>% 
  ggplot() +
    geom_textured_bar(aes(x = reorder(title, num_users), y = num_users, image = store), stat = "identity", width = 0.8) +
    scale_image_manual(values = images_fill) +
    labs(x = "", y = "Number of users (log scale)") +
    scale_y_continuous(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
    scale_y_log10(labels=function(x) format(x, big.mark = ",", scientific = FALSE)) +
    guides(image = "none") +
    theme_minimal() +
    theme(plot.margin = margin(0.6, 0, 0, 0, "cm")) +
    coord_flip()

chrome_combined <- plot_grid(chrome_user_bar, chrome_feature_plot, rel_widths = c(4,1), labels = c("Chrome Web", "Features"), hjust = "left", label_x = c(0.02, 0.075), label_y = 1)

save_plot(here::here("figures", "chrome_users_and_functionality.png"), chrome_combined,
          # each individual subplot should have an aspect ratio of 1.3
          base_aspect_ratio = 1.8
          )
```

## Combine them

```{r}
apple <- ggdraw() + draw_image(here::here("figures", "apple_ratings_and_features.png")) + theme(plot.margin = unit(c(0.2,0,0,0.2), "cm"))
play <- ggdraw() + draw_image(here::here("figures", "play_installs_and_functionality.png")) + theme(plot.margin = unit(c(0.2,0,0,0.2), "cm"))
chrome <- ggdraw() + draw_image(here::here("figures", "chrome_users_and_functionality.png")) + theme(plot.margin = unit(c(0.2,0,0,0.2), "cm"))
legend <- ggdraw() + draw_image(here::here("figures", "legend_horisontal.png"), scale = 0.8) + theme(plot.margin = unit(c(-8,0,0,0.5), "cm"))

user_numbers_combined_plot <- plot_grid(chrome, play, apple, legend, ncol=1)

save_plot(here::here("figures", "user_numbers_combined_plot.png"), user_numbers_combined_plot,
          ncol = 1, # we're saving a grid plot of 2 columns
          nrow = 4, # and 2 rows
          # each individual subplot should have an aspect ratio of 1.3
          base_aspect_ratio = 1.8
          )
```

# Plot of average ratings by design pattern types

```{r}
ratings_more_than <- 14

ave_ratings_table <- meta_data %>%
  filter(num_ratings > ratings_more_than) %>%
  group_by(categories_alphabetically) %>%
  summarise(num_tools = n(),
            median_rating = median(average_rating, na.rm = TRUE) %>% round(2),
            #mean_rating = mean(average_rating, na.rm = TRUE),
            min_rating = min(average_rating, na.rm = TRUE) %>% round(2),
            max_rating = max(average_rating, na.rm = TRUE) %>% round(2)) %>%
  arrange(desc(median_rating)) %>% 
  left_join(meta_data %>% distinct(categories_alphabetically, num_categories)) %>% 
  arrange(num_categories, desc(median_rating)) %>% 
  select(-num_categories) %>% 
  filter(!is.na(median_rating)) %>% 
  mutate(image_single = case_when(
    categories_alphabetically == "Block/removal" ~ "1_block",
    categories_alphabetically == "Goal-advancement" ~ "1_goal",
    categories_alphabetically == "Self-tracking" ~ "1_track",
    categories_alphabetically == "Reward/punish" ~ "1_reward",
    categories_alphabetically == "Block/removal + Goal-advancement" ~ "2_block_goal",
    categories_alphabetically == "Block/removal + Reward/punish" ~ "2_block_reward",
    categories_alphabetically == "Block/removal + Self-tracking" ~ "2_block_track",
    categories_alphabetically == "Goal-advancement + Reward/punish" ~ "2_goal_reward",
    categories_alphabetically == "Goal-advancement + Self-tracking" ~ "2_track_goal",
    categories_alphabetically == "Reward/punish + Self-tracking" ~ "2_track_reward",
    categories_alphabetically == "Block/removal + Goal-advancement + Self-tracking" ~ "3_block_track_goal",
    categories_alphabetically == "Block/removal + Reward/punish + Self-tracking" ~ "3_block_track_reward",
    categories_alphabetically == "Goal-advancement + Reward/punish + Self-tracking" ~ "3_goal_track_reward",
    categories_alphabetically == "Block/removal + Goal-advancement + Reward/punish + Self-tracking" ~ "4_all",
  )) %>% 
  mutate(image_single = str_c(here::here("figures", "plot_colours/"), image_single, ".png"))

ave_ratings_levels <- ave_ratings_table$categories_alphabetically

# now create the feature plot
images <- ave_ratings_table$image_single %>% 
  set_names(ave_ratings_table$categories_alphabetically)

feature_plot_compare_types <- ave_ratings_table %>% 
  ungroup() %>% 
  mutate(categories_alphabetically = factor(categories_alphabetically, levels = rev(ave_ratings_levels), ordered = TRUE)) %>% 
  ggplot() +
    geom_textured_bar(aes(x = categories_alphabetically, y = -0.9, image = categories_alphabetically), stat = "identity", width = 0.75) +
    scale_image_manual(values = images) +
    coord_flip() +
    labs(x = "", y = "") +
    guides(image = FALSE) +
    theme(axis.line=element_blank(),axis.text.x=element_blank(),
          axis.text.y=element_blank(),axis.ticks=element_blank(),
          axis.title.x=element_blank(),
          axis.title.y=element_blank(),legend.position="none",
          panel.background=element_blank(),panel.border=element_blank(),panel.grid.major=element_blank(),
          panel.grid.minor=element_blank(),plot.background=element_blank()) +
    theme(plot.margin = margin(1.5, 0, 0.82, 0, "cm"))

# create the boxplot
# with store
boxplot_ratings <- meta_data %>%
  filter(num_ratings > ratings_more_than) %>%
  mutate(categories_alphabetically = factor(categories_alphabetically, levels = rev(ave_ratings_levels), ordered = TRUE)) %>%
  rename(Store = store) %>% 
  ggplot(aes(x = categories_alphabetically, y = average_rating)) +
    geom_boxplot(outlier.shape = NA) +
    geom_jitter(width = 0.34, height = 0.06, size = 1.1, alpha = 0.8, aes(fill = Store, shape = Store)) +
    theme_minimal() +
    labs(x = "", y = "") +
    coord_flip() +
    theme(axis.text.y=element_blank(), plot.margin = margin(1.5, 0, 0, 0, "cm"), legend.position="left", legend.box.margin=margin(0,-0.5,0,0, "cm")) +
    scale_y_continuous(limits = c(1,5)) +
    scale_fill_manual(values = c("darkgrey", "#f0027f", "white")) +
    scale_shape_manual(values = c(22, 21, 25))

ratings_by_type <- plot_grid(boxplot_ratings, feature_plot_compare_types, rel_widths = c(3,1.2), labels = c("Average rating", "Features"), hjust = -0.14)

save_plot(here::here("figures", "ratings_by_type.png"), ratings_by_type,
          # each individual subplot should have an aspect ratio of 1.3
          base_aspect_ratio = 1.8
          )

ggdraw() + 
  draw_image(here::here("figures", "ratings_by_type.png")) +
  annotate(
    "text", label = "Block/\nremoval",
    x = 0.765, y = 0.84, size = 2.8, colour = "black"
  ) +
  annotate(
    "text", label = "Self-\ntracking",
    x = 0.83, y = 0.84, size = 2.8, colour = "black"
  ) +
  annotate(
    "text", label = "Goal-\nadvance",
    x = 0.895, y = 0.84, size = 2.8, colour = "black"
  ) +
  annotate(
    "text", label = "Reward/\npunish",
    x = 0.955, y = 0.84, size = 2.8, colour = "black"
  )

ggsave(here::here("figures", "ratings_by_type_annotated.pdf"), height = 4.3, width = 7)

```

# Wordcloud for reviews
```{r}
reviews <- read_csv(here::here("data", "processed", "distinct_reviews.csv"))

library(tidytext)
library(SnowballC)
library(wordcloud2)

data(stop_words)

tidy_review_words <- reviews %>% 
  unnest_tokens(word, review_text) %>% 
  anti_join(stop_words) %>% 
  filter(!is.na(word))

tidy_review_words %>%
  count(word, sort = TRUE) %>% 
  filter(!word %in% c("app", "apps", "phone", "extension")) %>% 
  wordcloud2(size = 1)

# tidy_review_words %>% 
#   mutate(word = case_when(
#     str_detect(word, "^love") ~ "love",
#     str_detect(word, "^block") ~ "block", 
#     str_detect(word, "^addict") ~ "addiction", 
#     str_detect(word, "^productiv") ~ "productivity", 
#     str_detect(word, "^wast") ~ "waste", 
#     str_detect(word, "track") ~ "tracking", 
#     str_detect(word, "^notification") ~ "notification", 
#     str_detect(word, "^feature") ~ "feature", 
#     str_detect(word, "^distract") ~ "distract", 
#     str_detect(word, "^focus") ~ "focus", 
#     word %in% c("recommend", "recommended") ~ "recommend", 
#     word == "use" ~ "use/usage", 
#     word == "usage" ~ "use/usage", 
#     TRUE ~ word
#   )) %>% 
#   count(word, sort = TRUE) %>% 
#   filter(!word %in% c("app", "apps", "phone", "extension")) %>% 
#   wordcloud2(size = 1)

# tidy_review_words %>% 
#   mutate(word = wordStem(word)) %>% 
#   count(word, sort = TRUE) %>% 
#   filter(!word %in% c("app", "apps", "phone", "extension")) %>% 
#   wordcloud2(size = 1)
```